Exploring Rural Shrink Smart Through Guided Discovery Dashboards

### 2022-06-06 SDSS Presentation/SDSS Presentation.Rmd
Updates from Brian adding break slides

<br><br>
.medium.emph[
Theme designed by Emily A. Robinson [www.emilyarobinson.com](https://www.emilyarobinson.com/)
]
???
Thank you for inviting me for a visit today and for attending my talk.

@@

<br>
.large.emph[
+ Introduction to Rural Shrink Smart Project with Iowa State University

+ Why Dashboards?

+ Dashboard Design and Guided Discovery Learning (GDL)

+ Parallel Coordinate Plots (PCPs) and the naive user

+ Data Collection and Feedback the loop

@@

@@

+ <span style="color:darkred">Project Question:</span> How do we leverage <span style="color:blue">data science</span>  and <span style="color:blue">exploratory data methods</span> to inform town leaders to reduce the shrinkage and maintain reliability?
    + How does an interactive dashboard of open-source data integrate into overall decision-making?
    + Can we focus on estimating only things that objective measurements don't capture?

# Identifying Key Metrics

<b>Audience</b>

.small.emph[
+ As small rural towns, traditionally, do not have access:
     + Data scientists 
     + Ability to easily leverage data collected locally to support decisions

Our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations 
]
--

<b>Visualization Goals</b>

.small.emph[
+ Create a central place for the excess of data to be conveyed in a single comprehensive visual.
    + We will utilize Parallel Coordinate Plot (PCPs)
+ Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g. Fire Dept, Schools, etc.)
+ Add an overview of town general basics related to the seven QoL factors.
]
--

<b>Engage People</b>

.small.emph[
+ We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
]

@@

+ How do we best design charts/graphs/UI to make people feel empowered and help them explore the data with an open mind?
    + How do we incorporate statistical analyses into this? Much of the data will provide qualitative differences but not statistically significant differences.

---
# Guiding Design Principles

@@

[Smart Connected Communities Dashbaord](https://shiny.srvanderplas.com/Smart-Connected-Communities/)
---
# Dashboard Philosophy

@@

--
<br>

.large.emph[
+ We allow users to modify a set of suggested comparison towns to indicate other towns of interest.
]
---
class:primary

@@

**Who?** Sukwon Lee, et al. wrote How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking

@@

NOVIS Model has been used to compare other graphs such as PCP, Tree Maps and Chord diagrams, which helps with the understanding of why we could use the PCPs. We will use this model to ask questions of our town leaders to gain feedback on what they understand.
---
# Dashboard Philosophy

@@

---
class:primary
# Dashboard Philosophy

@@

**Who?** [Michael A. DeDonno](https://doi.org/10.1016/j.lindif.2016.05.023) wrote The influence of IQ on pure discovery and guided discovery learning of a complex real-world task.

@@

@@

**Audience**

@@

## Dashbaord Feedback

@@

.center[
<img src="images/feedback_question.png" width="44%"/>
]
---
class:primary

@@

---
class:inverse
<br>
<br>
<br>
.center[
# Thank you! 
### Acknowledgments: Dr. Susan VanderPlas and the ISU - Rural Shrink Smart Project Team
<br>
<br>
`r fa("envelope", fill = "white")` **denise.bradford@huskers.unl.edu**
<!-- ##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/)  -->
`r fa("github", fill = "black")` **@drbradford12**
]

@@

Thank you for the support of my adviser Dr. Susan VanderPlas and the ISU Rural Shrink Smart

>>>>

???
Thank you for inviting me for a visit today and for attending my talk.

@@

<br>
.large.emph[
+ Introduction to Rural Shrink Smart Project with Iowa State University and Open-Source Data Challenges

+ Motivation - Why Dashboards?

+ Guiding Design Principles & Dashboard Design Workflow

+ Dashboard Philosophy - PCPs and the Novice User
  + Using Guided Discovery Learning (GDL)

+ Dashboard Design Workflow 
  + Feedback Loop (Active and Passive Methodologies)

@@

---
class:inverse
<br>
<br>
<br>
<br>
<br>
<br>
.center[
# Project Introduction and Open-source Data Description
]

@@

+ <span style="color:darkred">Project Question:</span> How do we leverage <span style="color:blue">data science</span>  and <span style="color:blue">exploratory data methods</span> to inform town leaders to reduce the shrinkage and maintain reliability?
    + How does an interactive dashboard of open-source data integrate into overall decision-making?
    + Can we focus on estimating only things that objective measurements don't capture?
---
class:inverse
<br>
<br>
<br>
<br>
<br>
<br>

.center[
# Motivation - Why Use Dashboards?
]
---

# Identifying Key Metrics
<b><u>Audience</u></b>

.medium.emph[
+ As small rural towns, traditionally, do not have access:
     + Data scientists 
     + Ability to easily leverage data collected locally to support decisions

Our research team will provide communities with data about services in small town Iowa in order to assist with developing strategies to improve quality of life for their residents amid shrinking populations 
]
--
<b><u>Visualization Goals</u></b>

.medium.emph[
+ Create a central place for the excess of data to be conveyed in a single comprehensive visual.
    + We will utilize Parallel Coordinate Plot (PCPs)
+ Create a map feature that will allow for town leaders to understand the distance of essential QoL components (e.g. Fire Dept, Schools, etc.)
+ Add an overview of town general basics related to the seven QoL factors.
]
--
<b><u>Engage People</u></b>

.medium.emph[
+ We will assess our visualizations to determine which strategies for user interface and interactive graphics design are most useful to empower town leaders to make discoveries in publicly available data assembled with a focus on items that impact rural quality of life.
]

@@

+ How do we best design charts/graphs/UI to make people feel empowered and help them explore the data with an open mind?
    + How do we incorporate statistical analyses into this? Much of the data will provide qualitative differences but not statistically significant differences.
---
class:inverse
<br>
<br>
<br>
<br>
<br>
<br>

.center[
# Guiding Design Principles and Dashboard Design Workflow
]    
---
# Guiding Design Principles

@@

[Smart Connected Communities Dashboard](https://shiny.srvanderplas.com/Smart-Connected-Communities/)
---
class:inverse
<br>
<br>
<br>
<br>
<br>
<br>
.center[
# Dashboard Philosophy - PCPs and the Novice User
]  
---
exclude: true

@@

--
exclude: true
<br>

.large.emph[
+ We allow users to modify a set of suggested comparison towns to indicate other towns of interest.
]

---
class:primary

@@

**Who?** Sukwon Lee, et al. wrote "How do People Make Sense of Unfamiliar Visualizations?: A Grounded Model of Novice’s Information Visualization Sensemaking"

@@

NOVIS Model has been used to compare other graphs such as PCP, Tree Maps and Chord diagrams, which helps with the understanding of why we could use the PCPs. We will use this model to ask questions of our town leaders to gain feedback on what they understand.

@@

---
class:primary

@@

**Who?** [Michael A. DeDonno](https://doi.org/10.1016/j.lindif.2016.05.023) wrote "The Influence of IQ on pure discovery and guided discovery learning of a complex real-world task."

@@

---
class:inverse
<br>
<br>
<br>
<br>
<br>
<br>
.center[
# Dashboard Design Workflow - Feedback Loop 
## Active & Passive Data Collection
]

@@

@@

## Dashboard Feedback

@@

.center[
<img src="images/feedback_question.png" width="55%"/>
]

---
class:primary

@@

???
With six rural town leaders, we will have enough information needed to understand the usefulness of the dashboard from

---
class:inverse
<br>
<br>
<br>
.center[
# Thank you! 
### Acknowledgments: Dr. Susan VanderPlas & the ISU - Rural Shrink Smart Project Team
<br>
<br>
`r fa("envelope", fill = "white")` **denise.bradford@huskers.unl.edu**
<!-- ##### `r fa("home", fill = "black")` [www.deniserbradford.com](https://www.deniserbradford.com/)  -->
`r fa("github", fill = "black")` **@drbradford12**
]

.medium.emph[
Special Thank You to Emily A. Robinson for creating our UNL theme, find her work at `r fa("github", fill = "black")` **@earobinson95**
]

@@

Thank you for the support of my adviser Dr. Susan VanderPlas and the ISU Rural Shrink Smart Project Team for the hard work that they have done for not only me but also the rural town leaders.
